---
title:
output:
  revealjs::revealjs_presentation:
    theme: white
    css: white-edits.css
    highlight: tango
    transition: slide
    self_contained: false
    reveal_plugins: ["chalkboard","search"]
    transitionSpead: fast
    mathjax: "https://cdnjs.cloudflare.com/ajax/libs/mathjax/2.7.1/MathJax.js?config=TeX-AMS_SVG"
    reveal_options:
      minScale: 1.0
      maxScale: 1.0
      width: 1024
      height: 768
      slideNumber: true
      controlsTutorial: true
      chalkboard:
        theme: whiteboard
        readOnly: false
---

## TITLE {data-background="../Backgrounds/title.jpg" class="header no-toc-progress" data-state="no-toc-progress"}

<!-- "reveal_plugins:..." and "chalkboard:" are handled by prep.py -->
```{r helpers, echo=FALSE, warning=FALSE, message=F}
library(knitr)
library(kableExtra)
html_df <- function(text, cols=NULL, col1=FALSE, full=F) {
  if(!length(cols)) {
    cols=colnames(text)
  }
  if(!col1) {
    kable(text,"html", col.names = cols, align = c("l",rep('c',length(cols)-1))) %>%
      kable_styling(bootstrap_options = c("striped","hover"), full_width=full)
  } else {
    kable(text,"html", col.names = cols, align = c("l",rep('c',length(cols)-1))) %>%
      kable_styling(bootstrap_options = c("striped","hover"), full_width=full) %>%
      column_spec(1,bold=T)
  }
}
```

<h3 class="date"><br>DATE</h3>
<h2 class="author"><br>Dr. Richard M. Crowley<br>rcrowley@smu.edu.sg<br><a href="http://rmc.link/">http://rmc.link/</a></h2>

# SECTION 1 {data-background="../Backgrounds/section.jpg" class="header"}

## BULLETS {data-background="../Backgrounds/default.jpg" class="default"}

- 1
    - 1.1 _italic_
- 2
    - 2.1 *bold*

## WEBSITE{data-background-iframe="http://quotes.wsj.com/index/SG/STI/advanced-chart"}

## QUOTE {data-background="../Backgrounds/default.jpg" class="default"}

> $Box1

## TWO COLUMN {data-background="../Backgrounds/default.jpg" class="default"}

<div class="clearfix">
<div class="left">
Left
</div>
<div class="right">
Right
</div>
</div>

## MATH {data-background="../Backgrounds/default.jpg" class="default"}

Some inline math like $\beta \times \gamma$ and:

$$
\begin{align*}
  \alpha &= 1 + 1 + 1\\
  &= 3
\end{align*}
$$

# SECTION 2 {data-background="../Backgrounds/section.jpg" class="header"}

## DATA: with percentage bars {data-background="../Backgrounds/default.jpg" class="default"}

```{r format_freq, echo=FALSE, warning = FALSE}
library(DT)
names <- c("Text", "Media", "Link", "Both")
vala <- as.vector(c(0.19,0.11,0.53,0.17))
valf <- as.vector(c(0.17,0.04,0.66,0.12))
valb <- as.vector(c(0.17,0.10,0.56,0.17))
valo <- as.vector(c(0.23,0.15,0.46,0.16))

df1 <- data.frame(Name=names, All=vala, Financial=valf, NonFinancial=valb, NonBusiness=valo)

datatable(df1, rownames = FALSE, options = list(
  searching = FALSE,
  columnDefs = list(list(width = '200px', targets = "_all")),
  fillContainer = TRUE,
  scrollY = 396,
  class = "display",
  paging = FALSE,
  info = FALSE
)) %>%
  formatPercentage(c('All','Financial','NonFinancial','NonBusiness'),digits=2) %>%
  formatStyle(
    names(df1[,-1]),
    background=styleColorBar(range(df1[,-1]),'#a2b1ff'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
  )
```

## DATA: frequency {data-background="../Backgrounds/default.jpg" class="default"}

```{r event_freq, echo=FALSE, warning = FALSE}
library(DT)
names <- c("M&A", "Dividends", "Financial", "Mgmt Forecast", "Executive","Awards", "Contract", "Analyst Forecast", "Insider trade", "10-K", "10-Q", "8-K")
vals <- as.vector(c(0.066, 0.025, 0.080, 0.027, 0.044, 0.0056, 0.032, 0.0090, 0.157, 0.0046, 0.013, 0.051))

df1 <- data.frame(Name=names, Percent=vals)

datatable(df1, rownames = FALSE, options = list(
  searching = FALSE,
  fillContainer = TRUE,
  scrollY = 396,
  class = "display",
  paging = FALSE,
  info = FALSE
)) %>%
  formatPercentage('Percent',digits=2) %>%
  formatStyle(
    'Percent',
    background=styleColorBar(df1$Percent,'#a2b1ff'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
  )
```

## DATA: plotly.js {data-background="../Backgrounds/default.jpg" class="default"}

```{r time graph, echo=FALSE, warning = FALSE, message = FALSE}
suppressMessages(library(plotly))
m <- list(
  l = 60,
  r = 50,
  b = 60,
  t = 40,
  pad = 4
)
suppressMessages(library(readr))

data <- read_csv("../Data/example.csv")
df = data.frame(data)

plot_ly(df,x=~days,y=~neutral, name='Neutral news', type='scatter', mode='lines') %>%
  add_trace(y=~negative, name='Negative news', mode='lines') %>%
  add_trace(y=~positive, name='Positive news', mode='lines') %>%
  layout(autosize = F, width = 800, height = 500, margin = m, title = "Tweets around financial news", 
         xaxis=list(title="Days relative to financial news"), yaxis=list(title="% days with tweets"))
```

## What is analytics? {data-background="../Backgrounds/default.jpg" class="default"}

> Simply put: Answering questions using data

- Additional layers we can add to the definition:
    - Answering questions using *a lot of* data
    - Answering questions using data *and statistics*
    - Answering questions using data *and computers*

```{r, eval=TRUE, include=FALSE, warning=FALSE, message=FALSE}
# Google ngrams for Data
library('ngramr')
ngd <- c("(Analytics + analytics)", "(Machine learning + machine learning)")
```
```{r, eval=TRUE, echo=FALSE, fig.width=7, fig.height=2.8}
suppressPackageStartupMessages(library(curl))
if(!is.null(curl::nslookup("r-project.org", error = FALSE))) {
  ggram(ngd, year_start=1960, geom = "area", google_theme=F, smoothing = 3) + theme(legend.position="bottom", legend.direction="horizontal")
} else {
  print("No internet to get data")
}
```
<div class="tag">Made using <a href="https://github.com/seancarmody/ngramr" class="ref">seancarmody/ngramr</a></div>

## GRAPH: gpplot2 + ggplotly {data-background="../Backgrounds/default.jpg" class="default"}

```{r graph1, echo=FALSE, warning=FALSE, message=FALSE, fig.height = 5, fig.width = 4.3}
library(readr)
df <- read_csv("../Data/example2.csv")
df <- df[order(df$tic,df$year),]
df$ni_lag <- c(NA, df$ni[1:nrow(df)-1])
df <- df[df$year == 2017,]
df <- df[!(is.na(df$ni) | is.na(df$ni_lag)),]
library(ggplot2)
suppressPackageStartupMessages(library(plotly))
cor <- cor(x=df$ni, y=df$ni_lag, method="pearson")
if(nrow(df)>200) {
  df <- df[sample(nrow(df), 200),]
}
```

<div class="left">
- Past company earnings predicts future company earnings
    - Some earnings are stable over time (Ohlsson model)
    - Correlation: `r cor`
</div>
<div class="right">
```{r graph1_part2, echo=FALSE, warning=FALSE, message=FALSE, fig.height = 5, fig.width = 4.3}
plot <- ggplot(df, aes(x=ni_lag, y=ni)) + 
  geom_point(shape=1, aes(text=sprintf("Ticker: %s", tic))) + 
  geom_smooth(aes(x=ni_lag, y=ni), method=lm, se=T) + 
  labs(x="2016 Net Income",
       y="2017 Net Income")
ggplotly(plot, tooltip="text")
```
</div>



## GRAPHS: replacing animation {data-background="../Backgrounds/default.jpg" class="default"}

```{r graph3, eval=TRUE, include=FALSE, warning=FALSE, message=FALSE}
drown <- c(109,102,102,98,85,95,96,98,123,94,102) # US CDC data, 1999 to 2009 (falling in)
drown2 <- c(421,465,494,538,430,530,511,600,582,605,603)
ic <- c(16.1, 15.8, 16.2, 15.9 ,14.6 ,15.1 ,15.3, 14.8, 14.2 ,13.9 ,14 ,13.2 ,13.2 ,13 ,12.5 ,12.9 ,13.1) # Statista, 2000 to 2016
icprice <- c(126.858,128.467,130.858,131.733,134.808,137.508,144.592,150.575,155.500,161.742,164.375,173.408,179.033,175.467,178.250,177.575,179.233,183.373,192.768,196.637,195.055,209.161,215.524,215.489,215.412,219.201,219.452,219.903,219.593) # CPI data for ice cream and related products, 1990-2018
cancerthursday <- c(80262,80994,81321,81988,81390,82682,82109,82516,82663,83166,84978,84742) # CDC (tylervigen.com)
# https://futurism.com/correlation-vs-causation-2/
df <- data.frame(drown=drown2[2:length(drown2)],
                 icecream=ic[1 : length(drown2)-1],
                 icecreamprice=icprice[11:(length(drown2)+9)],
                 year=2000:2009)
df$drown <- df$drown / mean(df$drown)
df$icecream <- df$icecream / mean(df$icecream)
df$icecream2 <- df$icecream*df$icecreamprice / mean(df$icecream*df$icecreamprice)
df$icecream3 <- df$icecreamprice / mean(df$icecreamprice)
cor_unit <- cor(x=df$icecream, y=df$drown)
cor_rev <- cor(x=df$icecream2, y=df$drown)
cor_price <- cor(x=df$icecreamprice, y=df$drown)
library(ggplot2)

library(reshape2)
dm <- melt(df[c("drown", "icecream2", "year")],id.var=3)
plot1 <- ggplot(dm, aes(x=year, y=value, colour=variable)) + 
  geom_point(shape=1) +
  geom_smooth(method=loess, se=F) +
  labs(x="Year",
       y="% of average year") + 
  scale_x_continuous(breaks=c(2000,2003,2006,2009)) + 
  scale_colour_discrete(name="Statistic",
                        breaks=c("icecream2","drown"),
                        labels=c("Ice cream revenue","Drowning in pools"))
dm <- melt(df[c("drown", "icecream3", "year")],id.var=3)
plot2 <- ggplot(dm, aes(x=year, y=value, colour=variable)) + 
  geom_point(shape=1) +
  geom_smooth(method=loess, se=F) +
  labs(x="Year",
       y="% of average year") + 
  scale_x_continuous(breaks=c(2000,2003,2006,2009)) + 
  scale_colour_discrete(name="Statistic",
                        breaks=c("icecream3","drown"),
                        labels=c("Ice cream volume sold","Drowning in pools"))
dm <- melt(df[c("drown", "icecream", "year")],id.var=3)
plot3 <- ggplot(dm, aes(x=year, y=value, colour=variable)) + 
  geom_point(shape=1) +
  geom_smooth(method=loess, se=F) +
  labs(x="Year",
       y="% of average year") + 
  scale_x_continuous(breaks=c(2000,2003,2006,2009)) + 
  scale_colour_discrete(name="Statistic",
                        breaks=c("icecream","drown"),
                        labels=c("Ice cream price","Drowning in pools"))
```

<div class="clearfix">
<div class="left">
- Ice cream revenue predicts pool drownings in the US
    - ???
    - Correlation is...<span class="fragment" data-fragment-index="1"> only `r cor_rev`</span>

<span class="fragment fade-up" data-fragment-index="2">
<ul><li>What about units sold?</li>
<ul><li>Correlation is negative!!!</li>
<li>`r cor_unit`</li></ul></ul>
</span>
<span class="fragment fade-up" data-fragment-index="3">
<ul><li>What about price?</li>
<ul><li>Correlation is `r cor_price`</li></ul></ul>
</span>
</div>
<div class="right fragment current-visible" style="position:relative; vertical-align: top;" data-fragment-index="1">
```{r graph3_part2, echo=FALSE, warning=FALSE, message=FALSE, fig.height = 4, fig.width = 4.3}
suppressPackageStartupMessages(library(plotly))
ggplotly(plot1)
```
</div>
<div class="right fragment current-visible" style="position:relative; vertical-align: top;" data-fragment-index="2">
```{r graph3_part4, echo=FALSE, warning=FALSE, message=FALSE, fig.height = 4, fig.width = 4.3}
suppressPackageStartupMessages(library(plotly))
ggplotly(plot3)
```
</div>
<div class="right fragment current-visible" style="position:relative; vertical-align: top;" data-fragment-index="3">
```{r graph3_part3, echo=FALSE, warning=FALSE, message=FALSE, fig.height = 4, fig.width = 4.3}
suppressPackageStartupMessages(library(plotly))
ggplotly(plot2)
```
</div>
</div>

> This is where the "educated" comes in

# SECTION: Coding {data-background="../Backgrounds/section.jpg" class="header"}

## What do individuals use analytics for? {data-background="../Backgrounds/default.jpg" class="default"}

- A great package for machine learning in python is <a href="https://radimrehurek.com/gensim/" class="ref">gensim</a>
- R:

```{r, eval=TRUE}
# Addition uses '+'
1 + 1

# Subtraction uses '-'
2 - 1
```

## Packages used for these slides {data-background="../Backgrounds/coding.jpg" class="default"}

- kableExtra
- knitr
- revealjs

## Custom code {data-background="../Backgrounds/coding.jpg" class="default"}

<small style="width:100%">
```{r, eval=F}
# List custom routines here

```
</small>

## NOTES {class="never"}

- Won't be displayed ever
